### 2014-08-20 Articles/repeated_equal_coeffs/literature.tex
Additional refactoring; add mtga styles (doesn't work well, however)

\begin{center}
СПИСОК ЛИТЕРАТУРЫ
\end{center}

\begin{enumerate}
\item
Самарский А.А., Гулин А.В. \textit{Численные методы}. М.: Наука, 1989.

\item
Гельфонд А.О. \textit{Исчисление конечных разностей}. М.: Книжный дом <<ЛИБРОКОМ>>, 2012.

\item
Aumann R.J., Maschler M. B. \textit{Repeated Games with Incomplete Information}. The MIT Press, Cambridge, London

\item
Domansky V. \textit{Repeated games with asymmetric information and random price fluctuations at finance markets} // Int J Game Theory. 2007. V. 36. P. 241--257.
\end{enumerate}