### 2014-08-21 Articles/repeated_equal_coeffs/footer.tex
Add header, introduction, complete model, and many other stuff

\begin{center}
СПИСОК ЛИТЕРАТУРЫ
\end{center}

\begin{enumerate}
\item
Самарский А.А., Гулин А.В. \textit{Численные методы}. М.: Наука, 1989

\item
Гельфонд А.О. \textit{Исчисление конечных разностей}. М.: Книжный дом <<ЛИБРОКОМ>>, 2012

\item
Aumann R.J., Maschler M. B. \textit{Repeated Games with Incomplete Information}. The MIT Press, Cambridge, London

\item
Domansky V. \textit{Repeated games with asymmetric information and random price fluctuations at finance markets} // Int J Game Theory. 2007. V. 36. P. 241--257

\item
De Meyer B., Saley H. \textit{On the stragegic origin of Brownian motion in finance} // Int J Game Theory. 2002. V. 31. P. 285--319

\item
Chatterjee K., Samuelson W. \textit{Bargaining under Incomplete Information} // Operations Research. 1983. V. 31. N. 5. P. 835--851
\end{enumerate}

\makeenginfo