### 2015-02-09 Articles/Convex Comb/bidding_convex.tex
Preparation for the second article

% -*- coding: cp866 -*-
%
% ****** maiksamp.tex 29.11.2001 ******
%
\documentclass[
aps,%
12pt,%
final,%
notitlepage,%
oneside,%
onecolumn,%
nobibnotes,%
nofootinbib,% 
superscriptaddress,%
noshowpacs,%
centertags]%
{revtex4}

\begin{document}
\selectlanguage{russian}

\title{НАЗВАНИЕ\\
СТАТЬИ}% Разбиение на строки осуществляется командой \\

\author{\firstname{А.~И.}~\surname{Пьяных}}
% Здесь разбиение на строки осуществляется автоматически или командой \\
\email{artem.pyanykh@gmail.com}
\affiliation{%
МГУ им. М.~В.~Ломоносова, ф-т ВМиК
}%

%\date{\today}
%\today печатает cегодняшнее число

\begin{abstract}
  Исследуется обобщение дискретной многошаговой модели биржевых торгов, рассмотренной в \cite{domansky}.
  Торги происходят между двумя игроками за однотипные акции, случайная цена акции может принимать два значения --- $ m $ с вероятностью $ p $ или $ 0 $ с вероятностью $ (1 - p) $ --- и определяется в начале торгов.
  Настоящая цена акции известна Игроку 1.
  Игрок 2 знает вероятность высокой цены акции и то, что Игрок 1 --- инсайдер.
  На каждом шаге торгов игроки делают целочисленные ставки.
  Игрок, предложивший большую ставку, покупает у второго акцию.
  Цена сделки определяется как выпуклая комбинация предложенных ставок.
  Данная модель сводится к повторяющейся игре с асимметричной информацией.
  Получено решение игры бесконечной продолжительности при произвольных значениях $ m $~и~$ p $: найдены оптимальные стратегии игроков и значение игры.
\end{abstract}

\maketitle

\section{Введение}

Some text. And another one.

And another one on the second line.

%
% Список литературы
%
\begin{thebibliography}{99}

\bibitem{gelfond} 
  \refitem{book} 
  Гельфонд А.О. \textit{Исчисление конечных разностей.} М.: Книжный дом <<ЛИБРОКОМ>>, 2012

\bibitem{kreps} 
  \refitem{article} 
  Крепс В.Л. \textit{Повторяющиеся игры, моделирующие биржевые торги, и возвратные последовательности.} // Изв. РАН. Теория и системы управления. 2009. Вып. 4. С. 109-120

\bibitem{samarsky}
  \refitem{book}
  Самарский А.А., Гулин А.В. \textit{Численные методы}. М.: Наука, 1989

\bibitem{sandomirskaya}
  \refitem{article}
  Сандомирская М.С., Доманский В.К. \textit{Решение одношаговой игры биржевых торгов с неполной информацией} // Математическая теория игр и её приложения. 2012. Т. 4. Вып. 1. С. 32-54

\bibitem{aumann}
  \refitem{article}
  Aumann R.J., Maschler M. B. \textit{Repeated Games with Incomplete Information}. The MIT Press, Cambridge, London

\bibitem{chatterjee}
  \refitem{article}
  Chatterjee K., Samuelson W. \textit{Bargaining under Incomplete Information} // Operations Research. 1983. V. 31. N. 5. P. 835--851

\bibitem{demeyer}
  \refitem{article}
  De Meyer B., Saley H. \textit{On the strategic origin of Brownian motion in finance} // Int J Game Theory. 2002. V. 31. P. 285--319

\bibitem{domansky}
  \refitem{article}
  Domansky V. \textit{Repeated games with asymmetric information and random price fluctuations at finance markets} // Int J Game Theory. 2007. V. 36(2). P. 241--257

\bibitem{myerson83}
  \refitem{article}
  Myerson R.B., Satterthwaite M.A. \textit{Efficient Mechanisms for Bilateral Trading} // Journal of Economic Theory. 1983. V. 29. P. 265-281
\end{thebibliography}
%
\newpage
%
% Для статей на русском языке далее следуют
% на английском языке название статьи, список авторов и краткая аннотация.
%
\selectlanguage{english}
\begin{center}
\large \bfseries \MakeTextUppercase{%
English name of paper
}
\end{center}
%
\begin{center}
\bfseries A.~Pyanykh
\end{center}
%
\begin{center}
\begin{minipage}{\textwidth - 2cm}
  \small%
  This paper is concerned with a modification of a discrete multistage
  bidding model introduced in \cite{domansky}. Bidding takes place
  between two players for one unit of a risky asset. The price of the
  asset is determined randomly at the start of the bidding and can be
  either $ m $ with a probability of $ p $ or $ 0 $ with a probability
  of $ (1 - p) $. The real price of the asset is known to Player
  1. Player 2 knows only the probability of a high price and that
  Player 1 is insider. At each stage of the bidding players make
  integral bids. The higher bid wins and one unit of the asset is
  transacted to the winning player. The price of the transaction
  equals to a convex combination of bids. This model is reduced to the
  repeated game with incomplete information. The solution for the
  infinite game with arbitrary $ m $ an $ p $ is found, including
  optimal strategies for both players and the value of the game.
\end{minipage}
\end{center}
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%% Character code reference %%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                             %
%     Upper case russian letters (CP866): АБВГДЕЁЖЗИЙКЛМНОПРСТУФХЦЧШЩЪЫЬЭЮЯ   %                                                                       %
%     Lower case russian letters (CP866): абвгдеёжзийклмнопрстуфхцчшщъыьэюя   %
%                     Upper case letters: ABCDEFGHIJKLMNOPQRSTUVWXYZ          %
%                     Lower case letters: abcdefghijklmnopqrstuvwxyz          %
%                                   Digits: 0123456789                        %
% Square, curly, angle braces, parentheses: [] {} <> ()                       %
%                Backslash, slash, solidus: \ / |                             %
%       Period, interrogative, exclamation: . ? !                             %
%                 Comma, colon, semi-colon: , : ;                             %
%          Underscore, hyphen, equals sign: _ - =                             %
%             Quotes (left, right, double): ` ' "                             %
%     Commercial-at, hash, dollar, percent: @ # $ %                           %
%  Ampersand, asterisk, plus, caret, tilde: & * +   ^                         %
%                                                                             %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% ****** End of file apssampr.tex ******